---
title: "Test Open Field Analysis"
author: "Adam Lu"
date: "December 7th, 2018"
output: html_document
---

## Preparation

Set parameters

```{r setup, include = FALSE}
# Print all code when knitting
knitr::opts_chunk$set(echo = TRUE)

# Define parameters
header <- c("Trial #", 
            "In Border:Periphery/Center-point:Duration", 
            "In Border:Periphery/Center-point:Frequency", 
            "In Border:Periphery/Center-point:Latency to first", 
            "In Center:Center:Duration", 
            "In Center:Center:Frequency", 
            "In Center:Center:Latency to first", 
            "In Zone:Center/Center-point:Duration", 
            "In Zone:Center/Center-point:Frequency", 
            "In Zone:Center/Center-point:Latency to first") 

# Old Code:
# header <- c("Trial #", "Animal ID", "Treatment", "Surgery", 
```

Load libraries

```{r, message = FALSE}
# tidyverse includes the packages readr, dplyr, ggplot2 ... etc.
library(tidyverse)

# ggbeeswarm includes geom_beeswarm()
library(ggbeeswarm)
```

## Read in the raw data

Read in the open field test raw data ([**Analysis Output-AV Open Field.csv**](Analysis Output-AV Open Field.csv)):

```{r, message = FALSE}
# Import open field data
openFieldDataRaw <- read_csv("Analysis Output-AV Open Field.csv", skip = 4, 
                            col_names = header, na = c("", "NA", "-"))

# Take a look at the raw open field data
openFieldDataRaw
```

Read in the data containing the treatment and sex information for each animal (**Injections Data.csv**):

```{r, message = FALSE}
# Import injections data
injectionsDataRaw <- read_csv("Injections Data.csv", n_max = 20)

# Take a look at the injections data
injectionsDataRaw
```

Read in a table that maps treatment numbers to treatment strings

```{r, message = FALSE}
# Create a dictionary for Treatment
treatmentTypeTable <- read_csv("Treatment Dictionary.csv")

# Display the table
treatmentTypeTable

# OLD CODE:
# treatmentTypeTable <- 
#     tibble(Treatment = 0:1, TreatmentStr = c("Control", "Ethosuximide"))
```

## Combine and rearrange the data

Starting from the open field raw data, do the following:
    1. Add the treatment and sex information by matching up the trial number
    2. Add the treatment string be matching up the treatment number
    3. Choose the column corresponding to duration in center and rename it
    4. Choose the column corresponding to duration in border and rename it
    5. Make "Sex"" a categorical variable
    6. Make "Treatment" a categorical variable
    7. Create a new column for the ratio of duration in border to duration in center

```{r, message = FALSE}
# Rearrange the data
openFieldDataExpanded <- 
    openFieldDataRaw %>% 
    full_join(injectionsDataRaw, by = c("Trial #" = "Trial # Open Field")) %>% 
    full_join(treatmentTypeTable, by = "Treatment") %>% 
    mutate(durationInCenter = `In Zone:Center/Center-point:Duration`) %>% 
    mutate(durationInBorder = `In Border:Periphery/Center-point:Duration`) %>% 
    mutate(Sex = factor(Sex)) %>% 
    mutate(TreatmentStr = factor(TreatmentStr)) %>% 
    mutate(durationRatio = durationInBorder / durationInCenter)

# Take a glimpse of all the variables in the data 
glimpse(openFieldDataExpanded)
```

Remove the first animal because it was used as a test animal

```{r, message = FALSE}
# Select all but the first row
openFieldDataFiltered <- 
    openFieldDataExpanded %>% 
    slice(2:n())

# Take a look at the data with rows filtered
openFieldDataFiltered
```

Select variables that will be analyzed

``` {r, message = FALSE}
# Select variables that will be analyzed
openFieldData <- 
    openFieldDataFiltered %>% 
    select(`Trial #`, TreatmentStr, Sex, durationInCenter, durationInBorder, durationRatio)

# Take a look at the data with columns filtered
openFieldData
```

## Make plots

#### Use Duration in Center
Plot the Duration in Center versus Treatment as a box plot

``` {r, message = FALSE}
# Add the data to the canvas
durationPlot <- ggplot(openFieldData, aes(TreatmentStr, durationInCenter)) +
    xlab("Treatment Group") + ylab("Duration in Center (s)") +
    ggtitle("Open Field Test for Ethosuximide-treated animals")

# Plot as a combination of jitter and box plot
durationPlot + geom_boxplot(outlier.color = 'red', alpha = 0.25) +
    geom_jitter(alpha = 0.25)

# Save the plot
ggsave("openField_durationCenter_boxplot_jitter.png")
```

Plot the Duration in Center versus Treatment as a violin plot

``` {r, message = FALSE}
# Plot as a combination of jitter and violin plot
durationPlot + geom_violin() +
    geom_jitter(alpha = 0.25)

# Save the plot
ggsave("openField_durationCenter_violin_jitter.png")
```

Plot the Duration in Center versus Treatment as a beeswarm plot

``` {r, message = FALSE}
# Plot as a combination of box and beeswarm plot
durationPlot + geom_boxplot(outlier.color = 'red', alpha = 0.25) +
    geom_beeswarm(alpha = 0.25)

# Save the plot
ggsave("openField_durationCenter_boxplot_beeswarm.png")
```
 
#### Use Ratio of Duration in Border to Duration in Center

``` {r, message = FALSE}
# Add the data to the canvas
durationRatioPlot <- ggplot(openFieldData, aes(TreatmentStr, durationRatio)) +
    xlab("Treatment Group") + ylab("Duration in Border/Duration in Center") +
    ggtitle("Open Field Test for Ethosuximide-treated animals")

# Plot as a combination of jitter and box plot
durationRatioPlot + geom_boxplot(outlier.color = 'red', alpha = 0.25) +
    geom_jitter(alpha = 0.25)

# Save the plot
ggsave("openField_durationRatio_boxplot_jitter.png")
```

Plot the Ratio of Duration in Border to Duration in Center as a violin plot

``` {r, message = FALSE}
# Plot as a combination of jitter and violin plot
durationRatioPlot + geom_violin() +
    geom_jitter(alpha = 0.25)

# Save the plot
ggsave("openField_durationRatio_violin_jitter.png")
```

Plot the Ratio of Duration in Border to Duration in Center as a beeswarm plot

``` {r, message = FALSE}
# Plot as a combination of beeswarm and box plot
durationRatioPlot + geom_boxplot(outlier.color = 'red', alpha = 0.25) +
    geom_beeswarm(alpha = 0.25)

# Save the plot
ggsave("openField_durationRatio_boxplot_beeswarm.png")
```











