---
title: "Test Novel Object Analysis"
author: "Alexa & Viktoria & Adam"
date: "December 10th, 2018"
output: html_document
---

```{r setup, include = FALSE}
# Print all code
knitr::opts_chunk$set(echo = TRUE)

# Define parameters
header <- c("Object Type", "Trial # NO", "Trial # OF", 
            "Novel Object", "Familiar Object", 
            "Distance to objects:object 1/Nose-point:Mean",
            "Distance to objects:object 2/Nose-point:Mean", 
            "Nose touching objects:object 1/Nose-point:Duration", 
            "Nose touching objects:object 1/Nose-point:Frequency", 
            "Nose touching objects:object 1/Nose-point:Latency to first",
            "Nose touching objects:object 2/Nose-point:Duration", 
            "Nose touching objects:object 2/Nose-point:Frequency", 
            "Nose touching objects:object 2/Nose-point:Latency to first")

```

Load libraries

```{r, message = FALSE}
# tidyverse includes the packages readr, dplyr, ggPlot ... etc.
library(tidyverse)

# ggbeeswarm includes geom_beeswarm()
library(ggbeeswarm)

# broom includes tidy()
library(broom)
```

## Read in the raw data

Read in the novel object test raw data ([**Analysis Output-VA Novel Object.csv**](Analysis Output-VA Novel Object.csv)):

```{r, message = FALSE}
# Import novel object data
novelObjectRaw <- read_csv("Analysis Output-VA Novel Object.csv", skip = 4, 
                            col_names = header)
# Look at Novel Object data
novelObjectRaw

```

Read in the data containing the treatment and sex information for each animal (**Injections Data.csv**):

```{r, message = FALSE}
# Import injections data
injectionsDataRaw <- read_csv("Injections Data.csv", n_max = 20)

# Take a look at the injections data
injectionsDataRaw
```

Read in a table that maps treatment numbers to treatment strings

```{r, message = FALSE}
# Create a dictionary for Treatment
treatmentTypeTable <- read_csv("Treatment Dictionary.csv")

# Display the table
treatmentTypeTable
```

## Combine and rearrange the data

Starting from the open field raw data, do the following: 
(TODO: UPDATE)  
    1. Add the treatment and sex information by matching up the trial number  
    2. Add the treatment string be matching up the treatment number  
    3. Choose the column corresponding to duration at object 1 and rename it  
    4. Choose the column corresponding to duration at object 2 and rename it  
    5. Make "Sex"" a categorical variable  
    6. Make "TreatmentStr" a categorical variable  
    6. Make "Object Type" a categorical variable and rename as "TrialType"  
    7. Compute the "DurationObjectRatio" variable  
    8. Compute the "DiscriminationIndex" variable (defined by duration at object 1 - duration at object 2 / (duration at object 1 + duration at object 2))  
    9. Create a "TrialTypeTreatment" variable that combines "TreatmentStr" and "TrialType"  

```{r, message = FALSE}
# Rearrange the data
novelObjectExpanded <- 
    novelObjectRaw %>% 
    full_join(injectionsDataRaw, by = c("Trial # OF" = "Trial # Open Field")) %>% 
    full_join(treatmentTypeTable, by = "Treatment") %>% 
    mutate(DurationObject1 = `Nose touching objects:object 1/Nose-point:Duration`) %>% 
    mutate(DurationObject2 = `Nose touching objects:object 2/Nose-point:Duration`) %>%
    mutate(Sex = factor(Sex)) %>%
    mutate(TreatmentStr = factor(TreatmentStr)) %>% 
    mutate(TrialType = factor(`Object Type`)) %>% 
    mutate(DurationObjectRatio = DurationObject1 / DurationObject2) %>% 
    mutate(DiscriminationIndex = (DurationObject1 - DurationObject2) / (DurationObject1 + DurationObject2)) %>% 
    mutate(TrialTypeTreatment = interaction(TreatmentStr, TrialType, sep = "-"))

# Take a glimpse of all the variables in the data 
glimpse(novelObjectExpanded)
```

Remove the first animal because it was used as a test animal:

```{r, message = FALSE}
# Select all but the first row
novelObjectFiltered <- 
    novelObjectExpanded %>% 
    slice(2:n())

# Take a look at the data with rows filtered
novelObjectFiltered

```

Select only the variables that will be analyzed:

``` {r, message = FALSE}
# Select variables that will be analyzed
novelObjectData <- 
    novelObjectFiltered %>% 
    select(TrialType, Treatment = TreatmentStr, Sex, 
           DurationObject1, DurationObject2, DurationObjectRatio, DiscriminationIndex, Tag, TrialTypeTreatment)

# Take a look at the data with columns filtered
novelObjectData

```

## Make plots

#### Use the Duration Object Ratio (Duration at Object 1 / Duration at Object 2)

Plot the Duration Object Ratio versus Trial Type as a box plot, colored by Treatment

``` {r, message = FALSE}
# Add the data to the canvas
durationRatioPlot <- 
    ggplot(novelObjectData, aes(TrialType, DurationObjectRatio, color = Treatment)) +
    labs(x = "Trial Type", y = "Duration Object Ratio", 
         title = "Novel Object Test for Ethosuximide-treated animals")

# Plot as a combination of jitter and box plot
durationRatioPlot + geom_boxplot(outlier.alpha = 0) +
    geom_point(alpha = 0.75, position = position_jitterdodge(jitter.width = 0.3))

# Save the plot
ggsave("novelObject_durationObjectRatio_boxplot_jitter.png")
```

Plot the Duration Object Ratio versus Trial Type as a violin plot

``` {r, message = FALSE}
# Plot as a combination of jitter and violin plot
durationRatioPlot + geom_violin() +
    geom_point(alpha = 0.75, position = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.9))

# Save the plot
ggsave("novelObject_durationObjectRatio_violin_jitter.png")
```

Plot the Duration Object Ratio versus Trial Type as a beeswarm plot

``` {r, message = FALSE}
# Plot as a combination of beeswarm and box plot
durationRatioPlot + geom_boxplot(outlier.alpha = 0) +
    geom_beeswarm(alpha = 0.75, dodge.width = 0.75)

# Save the plot
ggsave("novelObject_durationObjectRatio_boxplot_beeswarm.png")
```

#### Use Discrimination Index (duration at object 1 - duration at object 2 / (duration at object 1 + duration at object 2)

Plot the Discrimination Index versus Trial Type as a box plot, colored by Treatment

``` {r, message = FALSE}
# Add the data to the canvas
discriminationIndexPlot <- 
    ggplot(novelObjectData, aes(TrialType, DiscriminationIndex, color = Treatment)) +
    labs(x = "Trial Type", y = "Discrimination Index", 
         title = "Novel Object Test for Ethosuximide-treated animals")

# Plot as a combination of jitter and box plot
discriminationIndexPlot + geom_boxplot(outlier.alpha = 0) +
    geom_point(alpha = 0.75, position = position_jitterdodge(jitter.width = 0.3))

# Save the plot
ggsave("novelObject_discriminationIndex_boxplot_jitter.png")
```

Plot the Discrimination Index versus Trial Type as a violin plot

``` {r, message = FALSE}
# Plot as a combination of jitter and violin plot
discriminationIndexPlot + geom_violin() +
    geom_point(alpha = 0.75, position = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.9))

# Save the plot
ggsave("novelObject_discriminationIndex_violin_jitter.png")
```

Plot the Discrimination Index versus Trial Type as a beeswarm plot

``` {r, message = FALSE}
# Plot as a combination of beeswarm and box plot
discriminationIndexPlot + geom_boxplot(outlier.alpha = 0) +
    geom_beeswarm(alpha = 0.75, dodge.width = 0.75)

# Save the plot
ggsave("novelObject_discriminationIndex_boxplot_beeswarm.png")
```

## Perform statistics

#### Use the Duration Object Ratio (Duration at Object 1 / Duration at Object 2)

Use the Shapiro-Wilk test to test for normality

``` {r, message = FALSE}
# Test the entire data set for normality
shapiro.test(novelObjectData$DurationObjectRatio)

# Test each data set for normality
tapply(novelObjectData$DurationObjectRatio, novelObjectData$TrialTypeTreatment, shapiro.test)
```

Since the p-values for grouped data are greater than 0.05, the data sets can be considered normally distributed.

First, we apply the Welch Two Sample t-test for each trial type to test for any significant difference between treatment groups.

``` {r, message = FALSE}
# Filter the data by trial type
familiarTrialData <- 
  novelObjectData %>% filter(TrialType == "Familiar")
novelTrialData <- 
  novelObjectData %>% filter(TrialType == "Novel")

# Apply the Welch Two Sample t-test for each trial
tTestResultFamiliar <-
    t.test(DurationObjectRatio ~ Treatment, data = familiarTrialData)
tTestResultNovel <-
    t.test(DurationObjectRatio ~ Treatment, data = novelTrialData)

tTestResultFamiliar
tTestResultNovel
```

Since the p value > 0.05, there isn't any significant difference between treatment groups.

Next, we apply the Welch Two Sample t-test for each treatment type to test for any significant difference between trials.

``` {r, message = FALSE}
# Filter the data by trial type
controlData <- 
  novelObjectData %>% filter(Treatment == "Control")
ethosuximideData <- 
  novelObjectData %>% filter(Treatment == "Ethosuximide")

# Apply the Welch Two Sample t-test for each trial
tTestResultControl <-
    t.test(DurationObjectRatio ~ TrialType, data = controlData)
tTestResultEthosuximide <-
    t.test(DurationObjectRatio ~ TrialType, data = ethosuximideData)

tTestResultControl
tTestResultEthosuximide
```

Since the p value > 0.05, there is not really any significant difference between trials either.

#### Use Discrimination Index (duration at object 1 - duration at object 2 / (duration at object 1 + duration at object 2)

Use the Shapiro-Wilk test to test for normality

``` {r, message = FALSE}
# Test the entire data set for normality
shapiro.test(novelObjectData$DiscriminationIndex)

# Test each data set for normality
tapply(novelObjectData$DiscriminationIndex, novelObjectData$TrialTypeTreatment, shapiro.test)
```

Since the p-values for grouped data are greater than 0.05, the data sets can be considered normally distributed.

First, we apply the Welch Two Sample t-test to test for any significant difference between treatment groups, for each trial type.

``` {r, message = FALSE}
# Filter the data by trial type
familiarTrialData <- 
  novelObjectData %>% filter(TrialType == "Familiar")
novelTrialData <- 
  novelObjectData %>% filter(TrialType == "Novel")

# Apply the Welch Two Sample t-test for each trial
tTestResultFamiliar <-
    t.test(DiscriminationIndex ~ Treatment, data = familiarTrialData)
tTestResultNovel <-
    t.test(DiscriminationIndex ~ Treatment, data = novelTrialData)

tTestResultFamiliar
tTestResultNovel
```

Since the p value > 0.05, there isn't any significant difference between treatment groups for each trial type.

Next, we apply the Welch Two Sample t-test to test for any significant difference between trial types, for each treatment group.

``` {r, message = FALSE}
# Filter the data by trial type
controlData <- 
  novelObjectData %>% filter(Treatment == "Control")
ethosuximideData <- 
  novelObjectData %>% filter(Treatment == "Ethosuximide")

# Apply the Welch Two Sample t-test for each trial
tTestResultControl <-
    t.test(DiscriminationIndex ~ TrialType, data = controlData)
tTestResultEthosuximide <-
    t.test(DiscriminationIndex ~ TrialType, data = ethosuximideData)

tTestResultControl
tTestResultEthosuximide
```

Since the p value > 0.05, there is not really any significant difference between trial types for each treatment group.

However, there is the possibility that there is a significant difference between trial types across all treatment groups, or a significant difference between treatments across all trial types. Let's do an ANOVA (or a linear model) to measure if there is any interaction between Discrimination Index and trial type or treatment, when the other variable is accounted for.

```{r, message = FALSE}
# Two Way Anova Interaction Effect
novelObjectAnova <- aov(DiscriminationIndex ~ TrialType + Treatment, 
                        data = novelObjectData) 

# Print a summary of the ANOVA table
summary(novelObjectAnova)

# In table format
tidy(novelObjectAnova)
```

```{r, message = FALSE}
# Try modeling DiscriminationIndex against TrialType and Treatment.
novelObjectFit <- lm(DiscriminationIndex ~ TrialType + Treatment, data = novelObjectData)

# See what the model looks like
summary(novelObjectFit)

# In table format
tidy(novelObjectFit)
```

It looks like there is a significant interaction between discrimination index and trial type when treatment is accounted for, but not the other way around.

In fact, one can do a simple t-test on trial type regardless of treatment, but we are not trying to do this here.

```{r, message = FALSE}
# Apply the Welch Two Sample t-test for TrialType
tTestTrialType <-
    t.test(DiscriminationIndex ~ TrialType, data = novelObjectData)

tTestTrialType

# Apply the Welch Two Sample t-test for Treatment
tTestTreatment <-
    t.test(DiscriminationIndex ~ Treatment, data = novelObjectData)
tTestTreatment
```

## Conclusion

There is a significant difference in the discrimination index between trial types across all treatment groups (p value = **0.005**), but not a significant difference between treatments across all trial types (p value = **0.7**).

```{r, message = FALSE, include = FALSE}

# OLD CODE:
# # Filter data only for Familiar Object Trials
# familiarObject <-
# filter(novelObjectFiltered, `Object Type` == "Familiar")
# 
# #Filter data only for Novel Object Trials
# novelObject <-
# filter(novelObjectFiltered, `Object Type` == "Novel")

# durationRatioPlot1 <- 
#     ggplot(novelObjectData, aes(TrialTypeTreatment, DurationObjectRatio, color = TreatmentStr)) +
#     xlab("Treatment-Trial Type") + ylab("Duration Object Ratio") +
#     ggtitle("Novel Object Test for Ethosuximide-treated animals")
# durationRatioPlot1 + geom_boxplot(outlier.color = 'purple') +
#     geom_jitter(alpha = 0.75, width = 0.2)

# xlab("Trial Type") + ylab("Duration Object Ratio") + 
# ggtitle("Novel Object Test for Ethosuximide-treated animals")

# mutate(TrialTypeTreatment = interaction(TrialType, TreatmentStr, sep = "-"))

# # Look at Discrimination Index
# novelObjectData$DiscriminationIndex

       
```

